---
title: "2_Misophonia_PPS"
author: "GiuliaBrix"
date: "2025-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyverse)
options(scipen = 999)
```

#Import final dataset
```{r, message=FALSE, warning=FALSE}
Final_PPS_df <- read_csv("Misophonia_PPS_data/Final_PPS_df.csv")
names(Final_PPS_df)
```


#Define variables type
```{r, message=FALSE, warning=FALSE}
Final_PPS_df$ID <- factor(Final_PPS_df$ID)
Final_PPS_df$Group <- factor(Final_PPS_df$Group)
Final_PPS_df$Stimulus <- factor(Final_PPS_df$Stimulus)
Final_PPS_df$condition <- factor(Final_PPS_df$condition)
Final_PPS_df$trigger.time <- factor(Final_PPS_df$trigger.time)
Final_PPS_df$reaction.time <- as.numeric(Final_PPS_df$reaction.time)
```

#Density Plot to look at the distribution + geompoint plot 
```{r, message=FALSE, warning=FALSE}
Final_PPS_df <- Final_PPS_df %>% 
  filter(!is.na(reaction.time)) %>%  # Remove rows with NA in reaction.time
  filter(trigger.time != "0")       # Remove rows where trigger.time equals 0

ggplot(Final_PPS_df, aes(x = trigger.time, y = reaction.time, color = Group)) +
  geom_point(alpha = 0.7, size = 2) +  
  facet_grid(Group ~ Stimulus) +       
  theme_bw()
```

# Reaction times Descriptives
```{r, message=FALSE, warning=FALSE}
#Description of condition == VT1, VT2, VT3, VT4, VT5 reaction.time by Group, Stimulus and trigger.time
Final_PPS_df %>%
  filter(str_starts(condition, "VT")) %>%  # Filter rows where condition starts with "VT"
  group_by(Group, Stimulus, trigger.time) %>%  # Group by Group, Stimulus, and trigger.time
  summarise(
    mean_reaction_time = mean(reaction.time, na.rm = TRUE),
    sd_reaction_time = sd(reaction.time, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(Group, Stimulus, trigger.time)  # Arrange by Group, Stimulus, and trigger.time
```


#Corrected RTs
##1. Calculate the mean reaction time (Unimodal_mean) for each participant (ID) for unimodal trials (T) - this represents the time they usually need to react to a tactile stimulus
```{r, message=FALSE, warning=FALSE}

Final <- Final_PPS_df

unimodal_mean <- Final %>%
  filter(str_starts(condition, "T")) %>%  # Filter rows where condition starts with "T"
  group_by(ID) %>%  # Group by ID and trigger.time
  summarise(Unimodal_mean = mean(reaction.time, na.rm = TRUE), .groups = "drop")  # Calculate mean

```


##2.Corrected reaction times
Subtract to each reaction.time in bimodal trials (VT1, VT2, VT3, VT4, VT5) the Unimodal_mean for each participant.

```{r, message=FALSE, warning=FALSE}}
Final_b <- Final %>%
  left_join(unimodal_mean, by = c("ID")) %>%  
  mutate(diff_reaction_time = reaction.time - Unimodal_mean)  


ggplot(Final_b, aes(x = trigger.time, y = diff_reaction_time, color = Group)) +
  geom_point(alpha = 0.7, size = 2) +  
  facet_grid(Group ~ Stimulus) +       
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.5) +
  theme_bw()

#Export Final_b
#write_csv(Final_b, "Misophonia_PPS_data/Final_b.csv")
```

##3. Corrected RTs descriptives 
```{r, message=FALSE, warning=FALSE}
#Description of condition == VT1, VT2, VT3, VT4, VT5 diff_reaction_time by Group, Stimulus and trigger.time
Final_b %>%
  filter(str_starts(condition, "VT")) %>%  # Filter rows where condition starts with "VT"
  group_by(Group, Stimulus, trigger.time) %>%  # Group by Group, Stimulus, and trigger.time
  summarise(
    mean_diff_reaction_time = mean(diff_reaction_time, na.rm = TRUE),
    sd_diff_reaction_time = sd(diff_reaction_time, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(Group, Stimulus, trigger.time)  # Arrange by Group, Stimulus, and trigger.time

```

##Linear fitting
```{r, message=FALSE, warning=FALSE}
# Calculate mean difference in reaction time for each participant based on group, stimulus, and condition
RT2 <- Final_b %>%
  group_by(ID, Group, Stimulus, condition) %>%
  summarise(diff_reaction_time = mean(diff_reaction_time))

# Create wide format for analysis
RT2_wide <- RT2 %>%
  spread(condition, diff_reaction_time)

# Create empty columns for linear fit parameters
RT2_wide$slope <- NA
RT2_wide$intercept <- NA
RT2_wide$r_squared <- NA

# Perform linear fits for each participant
for (s in 1:nrow(RT2_wide)) {
  x = c(0.5, 1, 1.5, 2, 2.5)  # trigger times for VT1-VT5
  y = as.numeric(unlist(RT2_wide[s, c("VT1", "VT2", "VT3", "VT4", "VT5")]))
  
  # Fit linear model
  f <- lm(y ~ x)
  
  # Store parameters
  RT2_wide$slope[s] <- coef(f)[2]  # k in your equation
  RT2_wide$intercept[s] <- coef(f)[1]  # y0 in your equation
  RT2_wide$r_squared[s] <- summary(f)$r.squared
}

# Analyze slopes by group
group_summary <- RT2_wide %>%
  group_by(Group, Stimulus) %>%
  summarise(
    mean_slope = mean(slope, na.rm = TRUE),
    sd_slope = sd(slope, na.rm = TRUE),
    mean_intercept = mean(intercept, na.rm = TRUE),
    sd_intercept = sd(intercept, na.rm = TRUE),
    mean_r_squared = mean(r_squared, na.rm = TRUE)
  )

#Export group_summary
#write_csv(group_summary, "Misophonia_PPS_data/group_summary.csv")

ggplot(group_summary, aes(x = Stimulus, y = mean_slope, group = Group, color = Group)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +  # Points for mean slopes
  geom_line(position = position_dodge(width = 0.2)) +           # Lines connecting points within groups
  geom_errorbar(
    aes(ymin = mean_slope - sd_slope, ymax = mean_slope + sd_slope), 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  theme_bw() +
  labs(
    title = "Mean Slopes by Stimulus and Group",
    x = "Stimulus",
    y = "Mean Slope (k)",
    color = "Group"
  ) +
  theme(
    legend.position = "top",
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

```{r, message=FALSE, warning=FALSE}
group_fits <- RT2_wide %>%
  group_by(Group, Stimulus) %>%
  summarise(
    mean_slope = mean(slope, na.rm = TRUE),
    mean_intercept = mean(intercept, na.rm = TRUE)
  )

fit_lines <- group_fits %>%
  mutate(fit_line = purrr::map2(mean_slope, mean_intercept, ~ tibble(
    trigger_time = seq(0.5, 2.5, length.out = 100),
    reaction_time = .x * seq(0.5, 2.5, length.out = 100) + .y
  ))) %>%
  unnest(fit_line)

ggplot(fit_lines, aes(x = trigger_time, y = reaction_time, color = Stimulus)) +
  geom_line(size = 1.2) +  
  facet_wrap(~ Group) +  
  theme_bw() +
  labs(
    title = "Linear Fit",
    x = "Trigger Time",
    y = "Reaction Time",
    color = "Group"
  ) +
  theme(
    legend.position = "top",
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

```{r, message=FALSE, warning=FALSE}
ggplot(fit_lines, aes(x = trigger_time, y = reaction_time, 
                       color = Group, linetype = Stimulus)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("MS" = "red", "HC" = "blue"),
                     labels = c("MS" = "Misophonia", "HC" = "Control")) +
  scale_linetype_manual(values = c("NEUTRAL" = "dotted", "TRIGGER" = "solid")) +
  theme_bw() +
  labs(
    title = "",
    x = "Sound distance / Time point of approaching sound",
    y = "Corrected RTs",
    color = "Group",
    linetype = "Stimulus"
  ) +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )



```








